---
title: "Stat 291 - Recitation 5"
author: "Orçun Oltulu"
date: "26 / 11 / 2021"
output: 
  pdf_document: 
    latex_engine: xelatex
linkcolor: black
geometry: margin = 1.05in
fontsize: 12pt
---

```{r include = FALSE}
knitr::opts_chunk$set(echo=FALSE,results='hide',fig.show = 'hide')
```

# Last Week: 

Recall the last exercise from Recitation 4. 

Load 'Credit' data set from ISLR package. Read the document for the data set; '?Credit'.

```{r echo=TRUE}
library(ISLR)
library(dplyr)
```

### Part A.

Check the first 10 observations of Credit data set.

```{r}
head(Credit, 10)
```

### Part B.

Create a subset, new_credit1, for Asian married females. 

```{r}
new_credit1 <- filter(Credit, 
                      Gender == "Female", 
                      Married == "Yes", 
                      Ethnicity == "Asian")

head(new_credit1,10)
```

### Part C.

From new_credit1, select only numeric values and create new data frame, new_credit2.

```{r}
new_credit2 <- select_if(new_credit1, is.numeric)
head(new_credit2,10)
```

### Part D.

Find the mean of each numeric value in new_credit2.

```{r}
summarise_all(new_credit2, mean)
```

### Part F.

Now, find minimum and maximum Income for Asian males and females separately,

* Filter Asian people, 

* Select only Gender and Income variables, 

* Group them by Gender, 

* Summarize using min and max functions.

```{r}
new_credit3 <- filter(Credit, Ethnicity == "Asian")
new_credit3 <- select(new_credit3, Gender, Income)
new_credit3 <- group_by(new_credit3, Gender)

summarise(new_credit3, 
          Minimum_Income = min(Income), 
          Maximum_Income = max(Income)) 
```

# Reading data files into R:

## Exercise 1:

Read the file 'Table0.txt';

```{r}
df1 <- read.table("Table0.txt")
df1
```

**(a)** Assign names to the columns to Name, Age, Height, Weight and Sex.

```{r}
colnames(df1) <- c('Name', 'Age', 'Height', 'Weight','Sex');
df1
```

**(b)** Change the row names so that they are the same as Name, and remove the variable Name.

```{r}
row.names(df1) <- df1$Name
df1$Name <- NULL 
df1
```

## Exercise 2:

Read the file 'Table1.txt';

```{r}
df2 <- read.table("Table1.txt",header=T)
df2
```

**(a)** How many rows and columns does it have?

```{r}
ncol(df2)
nrow(df2)
```

**(b)** Reread the file and make the variable Name be the row names. Make sure you read the variable as characters and not as factors.

```{r}
df2.b <- read.table("Table1.txt", header=T, 
                    row.names = "Name", 
                    stringsAsFactors = FALSE)
df2.b

lapply(df2.b, class)
```


## Exercise 3:

Read the file 'Table2.txt';

```{r}
df3 <- read.table('Table2.txt', 
                  header = T, 
                  skip = 1) 
```

**(a)** What is the problem with the first and last columns ?

```{r results = TRUE}
df3
```

**(b)** How can you fix that problem ?

```{r}
df3.b <- read.table('Table2.txt', 
                  header = T, 
                  skip = 1, quote ="/") 
df3.b
```

## Exercise 4:

Read the file 'Table3.txt';

```{r}
df4 <- read.table('Table3.txt',header = T,skip = 1)
df4
```

**(a)** How many missing value does this data set have?

```{r}
sum(is.na(df4))
```

**(b)** Assign NA to each 'weird' value. 

```{r}
df4[3,2] <- df4[5,3] <- df4[6,4] <- NA
df4
```

**(c)** Reread the data set but this time make sure you only have 'NA' values rather than {'*','**', "--"}

```{r}
df4.c <- read.table('Table3.txt', header = T, skip = 1, 
                    na.strings = c("NA", "*", "**", "--"))
df4.c
```

## Exercise 5:

Read the file 'Table4.txt';

```{r}
df5 <- read.table('Table4.txt',header = T)
df5
```

Watch out for the missing values and the decimal separator.

```{r}
df5.a <- read.table('Table4.txt',header = T,
                    na.strings = c("NA", "*", "**","--"), 
                    dec = ",")
df5.a
```

## Exercise 6:

Read the file 'Table5.txt';

```{r}
df6 <- read.table('Table5.txt', header = T)
df6
```

Watch out for the missing values and the decimal separator and the separator.

```{r}
df6.a <- read.table('Table5.txt', header = T, 
                    na.strings = c(NA, "**", "--"),
                    dec = ",", sep = ";")
df6.a
```

## Exercise 7:

Read the file 'Table6.txt'; 

Check out the file first. Notice that the information is repeated, we only want the first non-repeated ones. Make sure to create only characters not factors this time around. Lastly, we don’t want the comments.

```{r}
df7.a<- read.table("Table6.txt", skip = 1,header = TRUE,
                   row.names = "Name",nrow = 7,
                   comment.char = "@",
                   stringsAsFactors = FALSE)
df7.a
```

## Exercise 8:

Read the file 'states1.csv';

```{r}
df8 <- read.csv("states1.csv")
head(df8,10)
```

**(a)** The names of the states should be the row names.

```{r}
df8.a <- read.csv("states1.csv",row.names = 1) 
head(df8.a,10)
```

**(b)** Check the dimensions of both 'df8' and 'df8.a' data.

```{r}
dim(df8);dim(df8.a)
```

## Exercise 9:

Read the file 'states2.csv';

```{r}
df9 <- read.csv("states2.csv",sep = ";") 
head(df9,10)
```

The names of the states should be the row names, watch out for the decimal separator and the separator.

```{r}
df9.a <- read.csv("states2.csv",row.names = 1,
                sep = ";",dec = ",") 
head(df9.a,10)
```

## Exercise 10:

Read the file 'states3.csv';

```{r}
df10 <- read.csv("states3.csv",row.names = 1,
                 sep = ";" ,dec = ",",
                 na.strings = c(NA, "*"))
head(df10,10)
```

Watch out for the same as the last exercise plus the missing values. Add to the previous data set, column-wise.

```{r}
head(cbind(df8.a, df9.a),10)
```


## Exercise 11:

Read 'iris.Rdata' into R.

```{r}
load("iris.Rdata")
```

# Writing data files:

## Exercise 12:

Using following commands create a data frame and write it to 'data1.txt' file. 

```{r echo=TRUE}
set.seed(291)
vec1 <- rnorm(15, mean = 5, sd = 2)
vec2 <- sample(100, size = 15)
vec3 <- runif(15)
vec4 <- sample(c("Male","Female"), 
               size = 15, replace = T)

data <- data.frame(vec1, vec2, vec3, vec4)
```

```{r}
write.table(data,file = "data1.txt")
```

## Exercise 13:

Write the same data set to 'data1.csv' file.  

```{r}
write.csv(data, "data1.csv")
```

&nbsp;

Add a new column c(15,25,0.5,"Female") to 'data1.csv' file.

```{r}
# first way
data.new <- rbind(data,c(15,25,0.5,"Female"))
write.csv(data.new, "data1.csv")
```

```{r}
# second way
write.table(data.frame(15,25,0.5,"Female"), "data1.csv", 
            append = TRUE,sep = ",", col.names=F)
```

## Exercise 14:

Create an .RData file with same data set, name it 'data1.Rdata'.

```{r}
save(data, file = "data1.Rdata")
```

# Reading Data from Web:

## Exercise 15: 

Read .csv data from from github. 

Go to the github repo with the following URL: 

https://github.com/oltuluorcun/Stat291

Read the "data1.csv" file in the repo. 

**Hint:** You have to use the **"Raw"** version of the data set and copy the raw versions URL.

```{r}
data_github <- read.csv("https://raw.github.com/oltuluorcun/Stat291/main/data1.csv")
data_github
```

## Exercise 16: 

**Web scraping with "rvest" package.** 

Go to the following web page where you will see the current F1 standings for both Drivers and Constructors. 

https://www.statsf1.com/en/2021.aspx

Then, using Selector Gagdet, find the 'location' of the tables. 

Finally, do the magic and extract the both tables and read into R.

```{r}
library(rvest)
webpage_standing <- "https://www.statsf1.com/en/2021.aspx"
webpage <- read_html(webpage_standing)
calendar_html <- html_nodes(webpage, ".yearclass")
    
drivers <- html_table(calendar_html[[1]], header = T, na.strings = "-")

drivers <- drivers[-grep("The drivers", drivers$Drivers),]

drivers <- data.frame(drivers)[c(2,25)]

constructors <- html_table(calendar_html[[2]], header = T, na.strings = "-")
    
constructors <- constructors[-grep("The constr", constructors$Constructors),]

constructors <- data.frame(constructors)[c(2,25)]

```

```{r}
head(drivers,10)
```


```{r}
constructors
```

&nbsp;

Check out for further info for usage of 'rvest' package: 

https://www.kaggle.com/orcunoltulu1/web-scraping-in-r-rvest-package


